---
title: "The Free Algebra in R"
output:
  github_document:
    pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
[![Total Downloads](https://cranlogs.r-pkg.org/badges/grand-total/freealg)](https://CRAN.R-project.org/package=freealg)
[![Build Status](https://travis-ci.org/RobinHankin/freealg.svg?branch=master)](https://travis-ci.org/RobinHankin/freealg)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/freealg)](https://cran.r-project.org/package=freealg)
[![Codecov test coverage](https://codecov.io/gh/RobinHankin/freealg/branch/master/graph/badge.svg)](https://codecov.io/gh/RobinHankin/freealg/branch/master)
<!-- badges: end -->

# Overview

The free algebra is an interesting and useful object.  Here I present
the `freealg` package which provides some functionality for free
algebra.

The package uses `C++`'s STL `map` class for efficiency, which has the
downside that the order of the terms is undefined.  This does not
matter as the mathematical value is unaffected by reordering; and the
print method does a good job in producing human-readable output.


# Installation

You can install the released version of `freealg` from
[CRAN](https://CRAN.R-project.org) with:

```{r, message=FALSE}
# install.packages("freealg")  # uncomment this to install the package
library("freealg")
```

# The free algebra

The free algebra is the free R-module with a basis consisting of all
words over an alphabet of symbols with multiplication of words defined
as concatenation.  Thus, with an alphabet of $\{x,y,z\}$ and

\[
A=\alpha x^2yx + \beta zy
\]

and 

\[
B=\gamma z + \delta y^4
\]

we would have

\[
A\cdot B=\left(\alpha x^2yx+\beta zy\right)\cdot\left(\gamma z+\delta y^4\right)=\alpha\gamma x^2yxz+\alpha\delta x^2yxy^4+\beta\gamma zyz+\beta\delta zy^5
\]

Note that multiplication is not commutative, but it is associative.  A
natural and easily implemented extension is to use upper-case symbols
to represent multiplicative inverses of the lower-case equivalents.
Thus if

\[
C=\epsilon X^2\]

we would have 

\[
A\cdot C=\left(\alpha x^2yx+\beta zy\right)\cdot\epsilon X^2=
\alpha\epsilon x^2yX + \beta\epsilon zyX^2
\]

and 

\[
C\cdot A=\epsilon X^2\cdot\left(\alpha x^2yx+\beta zy\right)=
\alpha\epsilon yx + \beta\epsilon X^2zy.
\]

The system inherits power associativity from distributivity and
associativity of concatenation, but is not commutative.

# The `freealg` package in use

Creating a free algebra object is straightforward.  We can coerce from
a character string with natural idiom:

```{r}
X <- as.freealg("1 + 3a + 5b + 5abba")
X
```
or use a more formal method:

```{r}
freealg(sapply(1:5,seq_len),1:5)
```

```{r}
Y <- as.freealg("6 - 4a +2aaab")
X+Y
X*Y
X^2
```

We can demonstrate associativity (which is non-trivial):

```{r}
set.seed(0)
(x1 <- rfalg(inc=TRUE))
(x2 <- rfalg(inc=TRUE))
(x3 <- rfalg(inc=TRUE))
```
(function `rfalg()` generates random `freealg` objects).  Then

```{r}
x1*(x2*x3) == (x1*x2)*x3
```


# Further information

For more detail, see the package vignette

`vignette("freealg")`
