---
title: "freealg_matrix"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Numerical verification of the freealg package

Here we verify that the `freealg` package is consistent.  

First, we observe that $n\times n$ matrices are an algebra over the
reals, and therefore should obey all the identities of a free algebra.
We will evaluate an expression in two different ways (one using the
package and the other using base R matrix algebra) and verify that the
results agree.  We will evaluate an expression on three
indeterminates, `A,B,C`, here random $5\times 5$ matrices:

```{r}
library("freealg")
 A <- matrix(rnorm(25),5,5)
 B <- matrix(rnorm(25),5,5)
 C <- matrix(rnorm(25),5,5)
```

Now we will take a random-ish element of the free algebra and call it
`x`:

```{r randfree}
x <- structure(list(indices = list(1L, 1:2, c(1L, 2L, 1L, 1L), 2L, 
    3L), coeffs = c(9, 6, 1, 9, 3)), class = "freealg")
x
```

We will evaluate `x^3` in two different ways.  Firstly, direct matrix
multiplication:

```{r}
xm <- 9*A + 6*A%*%B + A%*%B%*%A%*%A + 9*B + 3*C  # object "x" adapted for matrix multiplication
(xcubed_matrix <- xm %*% xm %*% xm)
```

And secondly using `freealg`:

```{r}
x^3
```

We can turn the above into an expression (rather clumsily) using
`sed`.  Cut-and-paste the above output into a file called `e.txt` [if
using the text above, strip out the `#` symbols and the first line
saying "free algebra..."] and then the following command

`cat e.txt|tr 'a-z' 'A-Z' | sed -e ':loop' -e 's/\([A-Z]\)\([A-Z]\)/\1%*%\2/g' -e 't loop'`

gives this:

```{r}
xcubed_freealg <- (
+ 729*A%*%A%*%A + 486*A%*%A%*%A%*%B + 81*A%*%A%*%A%*%B%*%A%*%A + 729*A%*%A%*%B + 486*A%*%A%*%B%*%A + 81*A%*%A%*%B%*%A%*%A%*%A + 54*A%*%A%*%B%*%A%*%A%*%A%*%B
+ 9*A%*%A%*%B%*%A%*%A%*%A%*%B%*%A%*%A + 81*A%*%A%*%B%*%A%*%A%*%B + 27*A%*%A%*%B%*%A%*%A%*%C + 324*A%*%A%*%B%*%A%*%B + 54*A%*%A%*%B%*%A%*%B%*%A%*%A + 486*A%*%A%*%B%*%B +
162*A%*%A%*%B%*%C + 243*A%*%A%*%C + 729*A%*%B%*%A + 486*A%*%B%*%A%*%A + 81*A%*%B%*%A%*%A%*%A%*%A + 54*A%*%B%*%A%*%A%*%A%*%A%*%B + 9*A%*%B%*%A%*%A%*%A%*%A%*%B%*%A%*%A
+ 81*A%*%B%*%A%*%A%*%A%*%B + 54*A%*%B%*%A%*%A%*%A%*%B%*%A + 9*A%*%B%*%A%*%A%*%A%*%B%*%A%*%A%*%A + 6*A%*%B%*%A%*%A%*%A%*%B%*%A%*%A%*%A%*%B + 1*A%*%B%*%A%*%A%*%A%*%B%*%A%*%A%*%A%*%B%*%A%*%A +
9*A%*%B%*%A%*%A%*%A%*%B%*%A%*%A%*%B + 3*A%*%B%*%A%*%A%*%A%*%B%*%A%*%A%*%C + 36*A%*%B%*%A%*%A%*%A%*%B%*%A%*%B + 6*A%*%B%*%A%*%A%*%A%*%B%*%A%*%B%*%A%*%A + 54*A%*%B%*%A%*%A%*%A%*%B%*%B +
18*A%*%B%*%A%*%A%*%A%*%B%*%C + 27*A%*%B%*%A%*%A%*%A%*%C + 324*A%*%B%*%A%*%A%*%B + 81*A%*%B%*%A%*%A%*%B%*%A + 54*A%*%B%*%A%*%A%*%B%*%A%*%A + 54*A%*%B%*%A%*%A%*%B%*%A%*%B +
9*A%*%B%*%A%*%A%*%B%*%A%*%B%*%A%*%A + 81*A%*%B%*%A%*%A%*%B%*%B + 27*A%*%B%*%A%*%A%*%B%*%C + 27*A%*%B%*%A%*%A%*%C%*%A + 18*A%*%B%*%A%*%A%*%C%*%A%*%B + 3*A%*%B%*%A%*%A%*%C%*%A%*%B%*%A%*%A +
27*A%*%B%*%A%*%A%*%C%*%B + 9*A%*%B%*%A%*%A%*%C%*%C + 972*A%*%B%*%A%*%B + 324*A%*%B%*%A%*%B%*%A + 81*A%*%B%*%A%*%B%*%A%*%A + 54*A%*%B%*%A%*%B%*%A%*%A%*%A +
36*A%*%B%*%A%*%B%*%A%*%A%*%A%*%B + 6*A%*%B%*%A%*%B%*%A%*%A%*%A%*%B%*%A%*%A + 54*A%*%B%*%A%*%B%*%A%*%A%*%B + 18*A%*%B%*%A%*%B%*%A%*%A%*%C + 216*A%*%B%*%A%*%B%*%A%*%B + 36*A%*%B%*%A%*%B%*%A%*%B%*%A%*%A
+ 324*A%*%B%*%A%*%B%*%B + 108*A%*%B%*%A%*%B%*%C + 162*A%*%B%*%A%*%C + 729*A%*%B%*%B + 486*A%*%B%*%B%*%A + 324*A%*%B%*%B%*%A%*%B +
54*A%*%B%*%B%*%A%*%B%*%A%*%A + 486*A%*%B%*%B%*%B + 162*A%*%B%*%B%*%C + 243*A%*%B%*%C + 162*A%*%B%*%C%*%A + 108*A%*%B%*%C%*%A%*%B + 18*A%*%B%*%C%*%A%*%B%*%A%*%A
+ 162*A%*%B%*%C%*%B + 54*A%*%B%*%C%*%C + 243*A%*%C%*%A + 162*A%*%C%*%A%*%B + 27*A%*%C%*%A%*%B%*%A%*%A + 243*A%*%C%*%B + 81*A%*%C%*%C +
729*B%*%A%*%A + 486*B%*%A%*%A%*%B + 81*B%*%A%*%A%*%B%*%A%*%A + 729*B%*%A%*%B + 486*B%*%A%*%B%*%A + 81*B%*%A%*%B%*%A%*%A%*%A + 54*B%*%A%*%B%*%A%*%A%*%A%*%B +
9*B%*%A%*%B%*%A%*%A%*%A%*%B%*%A%*%A + 81*B%*%A%*%B%*%A%*%A%*%B + 27*B%*%A%*%B%*%A%*%A%*%C + 324*B%*%A%*%B%*%A%*%B + 54*B%*%A%*%B%*%A%*%B%*%A%*%A + 486*B%*%A%*%B%*%B +
162*B%*%A%*%B%*%C + 243*B%*%A%*%C + 729*B%*%B%*%A + 486*B%*%B%*%A%*%B + 81*B%*%B%*%A%*%B%*%A%*%A + 729*B%*%B%*%B + 243*B%*%B%*%C +
243*B%*%C%*%A + 162*B%*%C%*%A%*%B + 27*B%*%C%*%A%*%B%*%A%*%A + 243*B%*%C%*%B + 81*B%*%C%*%C + 243*C%*%A%*%A + 162*C%*%A%*%A%*%B +
27*C%*%A%*%A%*%B%*%A%*%A + 243*C%*%A%*%B + 162*C%*%A%*%B%*%A + 27*C%*%A%*%B%*%A%*%A%*%A + 18*C%*%A%*%B%*%A%*%A%*%A%*%B + 3*C%*%A%*%B%*%A%*%A%*%A%*%B%*%A%*%A +
27*C%*%A%*%B%*%A%*%A%*%B + 9*C%*%A%*%B%*%A%*%A%*%C + 108*C%*%A%*%B%*%A%*%B + 18*C%*%A%*%B%*%A%*%B%*%A%*%A + 162*C%*%A%*%B%*%B + 54*C%*%A%*%B%*%C + 81*C%*%A%*%C +
243*C%*%B%*%A + 162*C%*%B%*%A%*%B + 27*C%*%B%*%A%*%B%*%A%*%A + 243*C%*%B%*%B + 81*C%*%B%*%C + 81*C%*%C%*%A + 54*C%*%C%*%A%*%B + 9*C%*%C%*%A%*%B%*%A%*%A
+ 81*C%*%C%*%B + 27*C%*%C%*%C
)
```
and then

```{r}
xcubed_freealg - xcubed_matrix
```

is zero, to within numerical tolerances.
